A Short Guide to Anyons and Modular Functors

### 2017-01-24 guide.tex
update after comment from B. Bartlett. Update refs.

\newcommand{\ket}[1]{|{#1}\rangle}
\newcommand{\expect}[1]{\langle{#1}\rangle}
\newcommand{\bra}[1]{\langle{#1}|}
\newcommand{\ketbra}[2]{\ket{#1}\!\bra{#2}}
\newcommand{\braket}[2]{\langle{#1}|{#2}\rangle}
\newcommand{\proj}[1]{\ketbra{#1}{#1}}

@@

\newcommand{\F}{\mathscr{H}} % H or F ?
\newcommand{\C}{\mathbb{C}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\Hom}{{\mathrm{Hom}}}

@@

Modular functors can be constructed from skein theory.
In the physics literature, 
this appears as skeins growing out of manifolds \cite{Pfeifer2012},
or as motivated by renormalization group considerations \cite{Levin2005}.
A further physical motivation is this: if a skein 
is supposed to correspond to the $(2+1)$-dimensional
world-lines of particles
as in the Schr\"{o}dinger picture,
modular functors would
correspond to the algebra of observables, as in
a Heisenberg picture.

In the physics literature, modular functors are
explicitly used in Refs.~\cite{Freedman2002, Freedman2002simulation}.
Also, Refs.~\cite{Beverland2014} and \cite{Kitaev2006topo}
use the language of modular functors but they call them TQFT's.
This is in fact reasonable because a modular functor can be
seen as part of a TQFT, but is quite confusing to the
novice who attempts to delve into the mathematical literature.

@@

Of the many variants of modular functor found in the literature,
one important distinction to be made is
the way anyons are labeled.
In the mathematical works \cite{Turaev1994, Bakalov2001, Tillmann1998} 
we see that anyons are allowed to have superpositions
of charge states. 
However, in this work we restrict anyons to have
definite charge states, as in \cite{Walker1991,Freedman2002simulation,Beverland2014}. 
This seems to be motivated physically as such configurations
would be more stable. % \simon{citation?}.

@@

We list the axioms for a 
\emph{2-dimensional unitary topological modular functor}.
 % abbreviated as \emph{MF}.
%There are quite alot of details that we leave out,
%%For the details of this construction, see
%see
%\cite{Turaev1994}, \cite{Walker1991} or \cite{Bakalov2001}.
%For a more pedestrian treatment, see \cite{Freedman2002simulation} or \cite{Beverland2014}.

@@

\begin{acknowledgments}

The author wishes to thank 
Andrew Doherty, Courtney Brell, Steven Flammia and Dominic Williamson
for helpful discussions and inspiration.

>>>>

% CUT HERE
\newcommand{\ket}[1]{|{#1}\rangle}
\newcommand{\expect}[1]{\langle{#1}\rangle}
\newcommand{\bra}[1]{\langle{#1}|}
\newcommand{\ketbra}[2]{\ket{#1}\!\bra{#2}}
\newcommand{\braket}[2]{\langle{#1}|{#2}\rangle}
\newcommand{\proj}[1]{\ketbra{#1}{#1}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\R}{\mathbb{R}}
% CUT HERE

@@

\newcommand{\F}{\mathscr{H}} % H or F ?
\newcommand{\A}{\mathcal{A}}
\newcommand{\Hom}{{\mathrm{Hom}}}

@@

Modular functors can be constructed from skein theory.
In the physics literature, 
this appears as skeins growing out of manifolds \cite{Pfeifer2012},
or as motivated by renormalization group considerations \cite{Levin2005}.
A further physical motivation is this: if a skein 
is supposed to correspond to the $(2+1)$-dimensional
world-lines of particles
as in the Schr\"{o}dinger picture,
the axioms for modular functors 
incorporate the algebra of observables, as in
a Heisenberg picture.
%modular functors would
%correspond to the algebra of observables, as in
%a Heisenberg picture.

In the physics literature, modular functors are
explicitly used in Refs.~\cite{Freedman2002, Freedman2002simulation}.
Also, Refs.~\cite{Beverland2016} and \cite{Kitaev2006topo}
use the language of modular functors but they call them TQFT's.
This is in fact reasonable because a modular functor can be
seen as part of a TQFT, but is quite confusing to the
novice who attempts to delve into the mathematical literature.

@@

Of the many variants of modular functor found in the literature,
one important distinction to be made is
the way anyons are labeled.
In the mathematical works \cite{Turaev1994, Bakalov2001, Tillmann1998} 
we see that anyons are allowed to have superpositions
of charge states. 
However, in this work we restrict anyons to have
definite charge states, as in \cite{Walker1991,Freedman2002simulation,Beverland2016}. 
This seems to be motivated physically as such configurations
would be more stable. % \simon{citation?}.

@@

We list the axioms for a 
\emph{2-dimensional unitary topological modular functor}.
 % abbreviated as \emph{MF}.
%There are quite alot of details that we leave out,
%%For the details of this construction, see
%see
%\cite{Turaev1994}, \cite{Walker1991} or \cite{Bakalov2001}.
%For a more pedestrian treatment, see \cite{Freedman2002simulation} or \cite{Beverland2016}.

@@

% CUT HERE
\begin{acknowledgments}

The author wishes to thank 
Andrew Doherty, Courtney Brell, Steven Flammia, Dominic Williamson and
Bruce Bartlett 
for helpful discussions and inspiration.
